---
title: "DATA-612 Project"
author: "Chen Hsu, Yen Chun Lin"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
    number_sections: TRUE
    includes:
    toc: true
    toc_float: true
    code_download: true
    theme: united
    df_print: kable
---

```{r setup, include=FALSE}
library(plotly)
library(dplyr)
library(tidyverse)
library(car)
```

```{r}
exit_velocity <- read.csv("exit_velocity.csv", header=T)
```

# Data
```{r}
head(n=10,exit_velocity)
```

## Variable information

- last_name..first_name: Players' name

- player_id: Player ID

- attempts: Batted ball events

- avg_hit_angle: Average hitting angle

- anglesweetspotpercent: Sweet Spot percentage (%)

- max_hit_speed: Max hitting speed (MPH)

- avg_hit_speed: Average hitting speed (MPH)

- gb: GB%

- max_distance: Max hitting distance (ft)

- avg_distance: Average hitting distance (ft)

- avg_hr_distance: Average home run hitting distance (ft)

- ev95plus: Exit Velocity

- ev95percent: Exit Velocity %

- barrels: Barrels

- brl_percent: Barrel %

- brl_pa: Barrel/PA


## Missing values
```{r}
show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  invisible(df)
}
show_missings(exit_velocity)
```

- There are no missing values in the initial data set.

## Rows, Columns
```{r}
dim(exit_velocity)
```

- There are 258 observations and 17 variables in the data set.

# Objective data
```{r}
bh<-exit_velocity%>%
  filter(max_distance>420)%>%
  select(last_name..first_name,player_id,max_distance,anglesweetspotpercent,avg_hit_angle,max_hit_speed)%>%
  filter(last_name..first_name %in% c("Olson, Matt","Schwarber, Kyle","	
García, Adolis","Alonso, Pete","	
Ohtani, Shohei","Acuña Jr., Ronald","	
Ozuna, Marcell","Riley, Austin","Seager, Corey","	
Betts, Mookie","	
Robert Jr., Luis","Alvarez, Yordan","Judge, Aaron","Muncy, Max","Soler, Jorge","	
Soto, Juan","Trout, Mike","Davis, J.D","Jiménez, Eloy","Correa, Carlos","Harris II, Michael","Pham, Tommy","Varsho, Daulton","Jones, Nolan","Bohm, Alec","Taylor, Michael A.","Murphy, Sean","Duvall, Adam","Bell, Josh","Suárez, Eugenio","Abreu, José","Langeliers, Shea","Steer, Spencer","	
Outman, James","Arozarena, Randy","Perez, Salvador","Reynolds, Bryan","Casas, Triston","Stanton, Giancarlo","Kepler, Max","Nimmo, Brandon","Ramírez, José","Alvarez, Francisco","Siri, Jose","Goldschmidt, Paul","Bellinger, Cody","Morel, Christopher","Arenado, Nolan","Suwinski, Jack","	
Guerrero Jr., Vladimir","Carroll, Corbin","Gorman, Nolan","Thomas, Lane","Turner, Trea","Witt Jr., Bobby","Henderson, Gunnar","Rooker, Brent","Raleigh, Cal","Machado, Manny","Santander, Anthony","Tucker, Kyle","Freeman, Freddie",""))%>%
  arrange(desc(max_distance))
head(bh,n=10)
```

- Setting 420 feet as the threshold because this is the home run distance that would result in a home run in all 30 stadiums in MLB.

- Choose the players who have more then ten home runs with max home run distance greater than 420 feet in 2023.


## Missing values
```{r}
show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  invisible(df)
}
show_missings(bh)
```

- There are no missing values in the objective data set.

## Rows, Columns
```{r}
dim(bh)
```

- There are 53 observations and 6 variables in the objective data set.

- There are 53 players who have more than ten home runs with max home run distance greater than 420 feet in 2023


## Diagnostic Plots
```{r}
hr<-lm(bh$max_distance~bh$anglesweetspotpercent+bh$max_hit_speed+bh$avg_hit_angle)
influenceIndexPlot(hr)
```

- No influential cases in the objective data set.


# Examples

## Plot 1
```{r}
sp<- bh %>%
  ggplot(aes(anglesweetspotpercent,max_distance , color = last_name..first_name)) + geom_point() + theme_bw()+ggtitle("Sweet Spot v.s. Max Home Run Distance")+labs(x="Sweet Spot Percentage", y="Max Home Run Disatance")

ggplotly(sp)
```

- The most optimal range of sweet spot percentage to have a home run over 420 feet is between 25 to 45 percent.

## Plot2
```{r}
a<- bh %>%
  ggplot(aes(x =avg_hit_angle,y = max_distance,color = last_name..first_name)) +geom_point() +theme_bw()+ggtitle("Hit Angle v.s. Max Home Run Distance")+labs(x="Hit Angle", y="Max Home Run Disatance")
ggplotly(a)
```

- The most optimal range of shitting angle to have a home run over 420 feet is between 5 to 30 degree.

## Plot3
```{r}
s<- bh %>%
  ggplot(aes(x =max_hit_speed , y=max_distance, color = last_name..first_name)) +geom_point() +theme_bw()+ggtitle("Max Hit Speed v.s. Max Home Run Distance")+labs(x="Max Hit Speed", y="Max Home Run Disatance")
ggplotly(s)
```

- The most optimal range of max hitting speed to have a home run over 420 feet is between 110 to 120 miles per hour.


# Conclusion

- In this data analysis, we utilized interactive plot to analyze and create three graphs illustrating hitting information in Major League Baseball (MLB) for the 2023 season. Each dot provides detailed information. 

- From the output, it is evident that the most optimal range for sweet spot percentage in hitting a home run over 420 feet is between 25% to 45%. Similarly, the most optimal hit angle for achieving a home run over 420 feet falls between 5 to 30 degrees. Additionally, the most effective range for the maximum hit speed in Miles Per Hour to secure a home run over 420 feet is between 110 to 120 MPH.


